### 2017-10-27 analysis/SEMwriteup.Rmd
ch.chh.chhh...changes

---
title: "Carnegie SEM"
author: "Sarah McKnight and Paul Harmon"
output: pdf_document
---

#Introduction:
We are attempting to run an SEM on the Carnegie Classification data to create a new classification system that does not rely on the PCA methodology currently in use. The Carnegie Classifications are currently based on two data-driven PCAs, meaning that loadings can change from year to year.  We hope to create a stable model that better allows for comparisons on a single dimension. This was done using Structural Equation Modeling (SEM).

We attempted to run an SEM on the original, raw data, but it was highly right-skewed, so we changed to using ranked data, as the Carnegie Classification currently uses.  Each of the metrics are ranked based on the minimum method for ties.  These data are much more uniformly distributed than the original data, which were highly skewed in the positive direction. These data included very large values for outlier institutions such as MIT and Harvard, but most of the schools were clustered at smaller values.

```{r, echo=FALSE}
#read in data
setwd("~/Carnegie-SEM/data")
cc2015 <- read.csv("CC2015data.csv",header = TRUE)

########2015################
cc2015.full <- read.csv("CC2015data.csv", header = TRUE, as.is = TRUE)
#updated file
#cc2015.full <- read.csv("Updated2015.csv", header = TRUE)

cc2015 <- cc2015.full[(cc2015.full$BASIC2015>14&cc2015.full$BASIC2015<18),]
cc2015$BASIC2015 <- factor(cc2015$BASIC2015)

#function for ranking the data
minrank <- function(x){rank(x, ties.method = "min")}

#dataset that we want to use
cc2015Ps<-
  na.omit(cc2015[,c("NAME","BASIC2010","BASIC2015","FACNUM","HUM_RSD","OTHER_RSD","SOCSC_RSD","STEM_RSD","PDNFRSTAFF","S.ER.D","NONS.ER.D")])

#calculate the ranked data
cc2015.r <- data.frame(cc2015Ps[,1:3],sapply(cc2015Ps[,-c(1:3)],minrank)) 
cc2015percap <- cc2015Ps[,c("PDNFRSTAFF","S.ER.D","NONS.ER.D")]/cc2015Ps$FACNUM
colnames(cc2015percap) <- c("PDNFRSTAFF_PC", "S.ER.D_PC", "NONS.ER.D_PC")
cc2015percap.r<-data.frame(sapply(cc2015percap,minrank))
cc2015_new <- cbind(cc2015Ps, cc2015percap)
cc2015_r <- cbind(cc2015.r, cc2015percap.r)
```

```{r data_plots, echo=FALSE}
par(mfrow=c(3,4))
hist(cc2015_r$FACNUM, xlab="FACNUM", main="")
hist(cc2015_r$HUM_RSD, xlab="HUM_RSD", main="")
hist(cc2015_r$OTHER_RSD, xlab="OTHER_RSD", main="")
hist(cc2015_r$SOCSC_RSD, xlab="SOCSC_RSD", main="")
hist(cc2015_r$STEM_RSD, xlab="STEM_RSD", main="")
hist(cc2015_r$PDNFRSTAFF, xlab="PDNFRSTAFF", main="")
hist(cc2015_r$S.ER.D, xlab="S.ER.D", main="")
hist(cc2015_r$NONS.ER.D, xlab="NONS.ER.D", main="")
hist(cc2015_r$PDNFRSTAFF_PC, xlab="PNDRFSTAFF_PC", main="")
hist(cc2015_r$S.ER.D_PC, xlab="S.ER.D_PC", main="")
hist(cc2015_r$NONS.ER.D_PC, xlab="NONS.ER.D_PC", main="")
```

We then attempted a regular SEM with the ranked data. This worked, but only with bootstrapped SE estimates and the fit was terrible (RMSEA = 0.9). Upon further exploration of the data, we realized that certain manifest variables in our data were highly correlated.

```{r, echo=FALSE}
cc2015_matrix2 <- as.matrix(cc2015_r[-c(1:3)])
corrmatrix <- Hmisc::rcorr(cc2015_matrix2)
corrplot::corrplot(corrmatrix$r, order="hclust")
```

So, with this new information in hand, we created a new model, aimed at minimizing RMSEA by modeling correlations. This model fit a lot better than the original model (RMSEA = 0.1) but the overall scores we got out were bimodal.

```{r}
model2 <- '
Aggregate=~HUM_RSD+OTHER_RSD+SOCSC_RSD+STEM_RSD+PDNFRSTAFF+S.ER.D+NONS.ER.D
PerCap=~PDNRSTAFF_PC+S.ER.D_PC+NONS.ER.D_PC

Overall=~Aggregate+PerCap

S.ER.D~~S.ER.D_PC
NONS.ER.D~~NONS.ER.D_PC
PDNFRSTAFF~~PDNRSTAFF_PC
PDNFRSTAFF~~S.ER.D
HUM_RSD~~SOCSC_RSD
S.ER.D~~STEM_RSD
PDNRSTAFF_PC~~S.ER.D_PC
NONS.ER.D~~HUM_RSD
NONS.ER.D~~SOCSC_RSD
'

lavaan_sem_r_cov <- lavaan::sem(model2, data=cc2015_r, std.lv=TRUE, orthogonal=FALSE, se="robust.huber.white")
semPlot::semPaths(lavaan_sem_r_cov)
#lavaan::summary(lavaan_sem_r_cov, fit.measures=TRUE)
CCScores_r_cov <- as.data.frame(lavaan::predict(lavaan_sem_r_cov))
density(CCScores_r_cov$Overall)
range_scores <- max(CCScores_r_cov$Overall) - min(CCScores_r_cov$Overall)
CCScores_r_cov$rate_2015 <- ifelse(CCScores_r_cov$Overall < min(CCScores_r_cov$Overall)+((1/3)*range_scores), 'A', ifelse(CCScores_r_cov$Overall > max(CCScores_r_cov$Overall)-((1/3)*range_scores), 'C', 'B'))
CC_table <- as.data.frame(cbind(cc2015_new$NAME, cc2015_new$BASIC2015, CCScores_r_cov$rate_2015))
colnames(CC_table) <- c("name", "basic2015", "rate2015")
table(cc2015_new$BASIC2015, CC_table$rate2015)
```

Recalling that our original data did not follow a modal distribution, the results seemed strange, and the contingency table showed that we only seemed to able to match up with the highest rating group. With that haing seemingly failed in producing three groups, we switched to Bayesian SEM models.

#Paul's Bayesian Stuff
An alternative to the traditional SEM model is the Bayesian SEM, which allows for prior distributions to be put on the latent variables.

In a traditional SEM model, the manifors

These models can be fit using the package "blavaan" (Merkle and Rosseel 2016); however, it is a beta package that is still being tested for validity and stability.

## Testing with Linear Regression
To better understand the mechanics of the R package Blavaan, we tried fitting a simple regression model to a dataset with which were already familiar. We were able to match the estimates with a linear model using lm(), which is expected. The choice of priors was Normal.

## A First Model

## Changing Priors

## Sensitivity to Priors

##Bayesian SEM - Some Conclusions

#Comparing the SEM model to the Carnegie Classifications
The SEM model and the PCA-based Carnegie model were both used to make plots of aggregate and per-capita traits associated with each institution. Neither plot produces multiple well-separated groups; however, the PCA-based method is slightly more well-separated.

```{r}
###lets see if we can calculate distances with schools
Score_dist <- as.data.frame(cbind(CCScores_r_cov_scale, PC_pred_scale))
rownames(Score_dist) <- cc2015_new$NAME

#recreates the PC-based plot
ggplot(Score_dist) + geom_point(aes(x = PC_ag_pred, y = -PC_percap_pred, color = cc2015_new$BASIC2015)) + ggtitle("Carnegie Classification PC Plot")

ggplot(Score_dist) + geom_point(aes(x = Aggregate, y = PerCap, color = CCScores_r_cov$rate_2015))

#calculate distances 
get_Distance <- function(x1,x2,y1,y2){

 #calculates the euclidean distance between the points  
  dist <- sqrt((x2 - x1)^2 + (y2 - y1)^2)
 #returns distances
  return(dist)
}

Score_dist$Distance <- c()

for (j in 1:nrow(Score_dist)){
  Score_dist$Distance[j] <- get_Distance(x1 = Score_dist[j,"Aggregate"],x2 = Score_dist[j,"PC_ag_pred"],y1 = Score_dist[j,"PerCap"],y2 = Score_dist[j,"PC_percap_pred"])
}

#Let's look at distances
sorted <- Score_dist[order(Score_dist$Distance),]
head(sorted,10)
tail(sorted,10)

summary(Score_dist$Distance)

#show biggest movers on Carnegie Classifications Plot
ggplot(Score_dist) + geom_point(aes(x = PC_ag_pred, y = PC_percap_pred, color = as.factor(round(Distance)))) + ggtitle("Biggest Movers")

```