### 2017-11-01 SEM_presentation.Rmd
presentation

---
title: "Carnegie SEM"
author: "Paul Harmon and Sarah McKnight"
date: "November 1, 2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The Carnegie Classifications

- Based on PCA
- Two Scales: Aggregate (x-axis) and Per-Capita (y-axis)

![The 2015 Classifications]()

## How are Scores Calculated?
The Carnegie Classifications are based on ranked data (because some schools are vastly larger on certain metrics than the rest). The ranked data are partitioned into two datasets and a PCA is generated on the both datasets. The first PC score is then used as an index for each trait.

## Aggregate and Per-Capita Indices

The classifications are calculated based on two indices of institutional output. The first is based on a weighted average of the number of PhDs awarded by the institution; the second is based on a per-capita measurement of research expenditures and research staff. 
__Aggregate Index:__
$$Ag.Index_{i}  = HumanitiesPhD_{i} + StemPhD_{i} + SocialSciencePhD_{i} + OtherPhD_{i}  + StemExpenditures_{i} + NonStemExpenditures_{i} + ResearchStaff_{i} $$ 
__Per Capita Index:__ 
$$PC.Index_{i} = \frac{ResearchStaff_{i} + StemExpenditures_{i} + NonStemExpenditures_{i}}{FacultySize_{i}} $$

## PCA Plot - Arbitrary Boundaries

From here, the methodology of the Carnegie Classifications is to produce a plot of the two correlated indices. Lines are arbitrarily drawn to create three partitions of the space. The following plot is produced.

![The 2015 Classifications with lines](https://github.com/paulharmongj/Writing-Project/Carnegie Classifications/Presentation_for_writing_maybe_files/figure-beamer/unstandardized scores-1.pdf)

## Problems with PCA
Many of the decisions made in this process are data driven and are outside the control of researchers:

+ Aggregate and Per-Capita Indices were constructed so that they are correlated, 
+ Attributes of the PCA change from year to year, meaning important policy drivers in one year might not be important in the next year (Social Science PhDs) 
+ Data are rarely grouped into three neat clusters, and lines are arbitrarily drawn

__Because of this, we wanted to try fitting a Structural Equation Model to create an index to compete with the Carnegie Classifications__

## Structural Equation Models
__GoaL__: Model two latent traits, one for the aggreagate scale and the other for the per-capita scale, and then use those to generate a "super-latent" single trait.

We could then separate institutions into groups via univariate model-based clustering or something as simple as breaking the single variable into several groups.

##SEM Model

##Path Diagram

##Questions