### 2018-04-05 for_laura.Rmd

---
title: "For Laura"
author: "Paul Harmon and Sarah McKnight"
date: "December 11, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We generated a new SEM model for the Carnegie Data in order to get things to run efficiently. The code for generating the model and reading in the dataset is given below.

```{r}
#read in data
setwd("~/Carnegie-SEM/data")
cc2015 <- read.csv("CC2015data.csv",header = TRUE)

########2015################
cc2015.full <- read.csv("CC2015data.csv", header = TRUE, as.is = TRUE)
#updated file
#cc2015.full <- read.csv("Updated2015.csv", header = TRUE)

cc2015 <- cc2015.full[(cc2015.full$BASIC2015>14&cc2015.full$BASIC2015<18),]
cc2015$BASIC2015 <- factor(cc2015$BASIC2015)

#function for ranking the data
minrank <- function(x){rank(x, ties.method = "min")}

#dataset that we want to use
cc2015Ps<-
  na.omit(cc2015[,c("NAME","BASIC2010","BASIC2015","FACNUM","HUM_RSD",
                    "OTHER_RSD","SOCSC_RSD","STEM_RSD","PDNFRSTAFF","S.ER.D","NONS.ER.D")])

#calculate the ranked data
cc2015.r <- data.frame(cc2015Ps[,1:3],sapply(cc2015Ps[,-c(1:3)],minrank))

cc2015percap <- cc2015Ps[,c("PDNFRSTAFF","S.ER.D","NONS.ER.D")]/cc2015Ps$FACNUM
colnames(cc2015percap) <- c("PDNRSTAFF_PC", "S.ER.D_PC", "NONS.ER.D_PC")
cc2015percap.r<-data.frame(sapply(cc2015percap,minrank))
cc2015_r <- cbind(cc2015.r, cc2015percap.r)

cc2015_matrix2 <- as.matrix(cc2015_r[-c(1:3)])
corrmatrix <- Hmisc::rcorr(cc2015_matrix2)
corrplot::corrplot(corrmatrix$r, order="hclust")
```

Based on the inability of the previous model to generate a positive-definite variance covariance matrix, we looked back at the model and determined that things are too highly correlated to get things to work. Rather than modeling the Aggregate and Per-Capita traits from the same set of variables, we decided to look at STEM vs. Non-STEM differences. This allowed us to have two correlated factors, but they did not depend on the same variables so we were able to get results that looked reasonable.

The model fit is below: 
```{r}
model4 <- '
HUMANITIES=~HUM_RSD+OTHER_RSD+SOCSC_RSD+NONS.ER.D+FACNUM
STEM=~STEM_RSD+PDNFRSTAFF+S.ER.D+FACNUM
Aggregate=~HUMANITIES+STEM'

lavaan_sem_new <- lavaan::sem(model4, data=cc2015_r, std.lv=TRUE,
                              orthogonal=FALSE, se="robust.huber.white")
lavaan::summary(lavaan_sem_new, standardized=TRUE, fit.measures=TRUE)
CCScores <- as.data.frame(lavaan::predict(lavaan_sem_new))
rownames(CCScores) <- cc2015Ps$NAME

```

The model would look like the following, in a path diagram:

```{r}
library(semPlot)
semPaths(lavaan_sem_new)
```

Finally, we can generate a plot of the factors plotted against each other: 
```{r}
library(ggplot2)
CCScores_r_cov_new <- as.data.frame(lavaan::predict(lavaan_sem_new))
names(CCScores_r_cov_new)
    #creates a plot and colors by Carnegie Classification Colors  
  ggplot(CCScores_r_cov_new) + geom_point(aes(x = STEM, y = HUMANITIES), color = "gray30", symbol = "gray30")+ 
   ggtitle("Predicted vs Actual Classifications") + theme_bw() + coord_fixed(ratio = 1)
```